---
title: "SIR Model Example"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(deSolve)
```

**Credit:** This example is built on (and closely follows) the example presented [here](http://rstudio-pubs-static.s3.amazonaws.com/6852_c59c5a2e8ea3456abbeb017185de603e.html) by Dr. Aidan Findlater

## Set up: define parameters

```{r}
### Specify parameters

## Proportion in each compartment: Susceptible 0.999999, Infected 0.000001, Recovered 0 (these are directly from the example linked above)
#these are the initial proportions of people who are inefected or susceptible
init <- c(S = 1-1e-6, I = 1e-6, R = 0.0)

## beta: infection parameter; gamma: recovery parameter

# original parameters <- c(beta = 2.316, gamma = 0.261)
parameters <- c(beta = 1.316, gamma = 2.261)
#increase the infection parameter, people get infected faster 
#increasing recovery parameter recovery happens faster
## Time frame
#if you decrease the inection paramter enough then no one gets infected 
times      <- seq(0, 100, by = 0.5)
```

## Create the function with the differential equations:

```{r}
## Build the function with all 
sir <- function(time, init, parameters) {

  with(as.list(c(init, parameters)), {

    dS <- -beta * S * I
    dI <-  beta * S * I - gamma * I
    dR <-  gamma * I
    
    return(list(c(dS, dI, dR)))
  })
}
```

## Approximate the solution using `deSolve::ode()`:
```{r}
## Solve using `deSolve::ode()`
approximation <- ode(y = init, times = times, func = sir, parms = parameters)
```

## Make it something we can plot:
```{r}
## Get output into a data frame
approx_df <- as.data.frame(approximation)

## Pivot longer so R will do the work for us: 
approx_long <- approx_df %>% 
  pivot_longer(cols = S:R, names_to = "population", values_to = "proportion")

```

## Plot the output with ggplot:
```{r}
ggplot(data = approx_long, aes(x = time, y = proportion)) +
  geom_line(aes(color = population))
```
Explore the code and output graph. What is the code doing? What is the graph telling us?
the graph is telling us the proportion of people that are susceptible is high and the proportion of people who have recovered is low. As people become infected less people are susceptible and more people are recovered. 
## End
